A New Hybrid Machine Learning Method for Stellar Parameter Inference

### 2025-03-05 paper/paper2/paper.tex
Nearly finished! Just need to cite like one more thing!

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is expected to be more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
    We first use \texttt{blase}'s nontraditional semi-empirical approach, \textbf{recasting} spectra \textbf{into} sets of tunable spectral lines.
    \texttt{blase} is \textbf{run} on 1,314 spectra from \textbf{a rectilinear subset of} the PHOENIX synthetic spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
    These manifolds are aggregated to create the PHOENIX generator, \textbf{enabling} parallelized reconstruction of spectra given stellar parameters.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square loss between input and PHOENIX generator spectra.
    \textbf{In} testing \textbf{210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex), we find inference errors of:} $T_{\mathrm{eff}}$\textbf{:} 185 K, $\log(g)$\textbf{:} 0.19, and [Fe/H]\textbf{:} 0.12 dex.
    \textbf{We also publish online} an archive of the \texttt{blase} \textbf{models} of the PHOENIX subset.
    This study is a proof of concept showing that semi-empirical HML is a viable alternative to traditional approaches.
\end{abstract}

@@

\begin{table*}
    \hspace*{0.6cm}\begin{tabular}{llll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Interval} & \bf{Sampling}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex & \bf{N/A}\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex & \bf{0.5 dex}\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
        Surface Gravity & $\log(g)$ & [2, 6] & \bf{0.5} \\
        Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study.
    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}.
    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}.
    This subset is comprised of 1,314 individual spectra: 73 $T_{\mathrm{eff}}$ values, 9 $\log(g)$ values, and 2 [Fe/H] values.}
    \label{tab:table1}
\end{table*}

@@

Spectral lines were often only detected in some spectra from the PHOENIX subset.
In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of the PHOENIX subset.
    We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    \textbf{Remember} that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ appear.
    Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}}
    \label{fig:figure2}
\end{figure*}

@@

Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66\AA \ C I and the 10123.66\AA \ He II lines \citep{NIST} varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12,000, $\log(g)$ = 6, [Fe/H] = -0.5) and the maximum being 34,551 ($T_\text{eff}$ = 2,300, $\log(g)$ = 3, [Fe/H] = 0).}
Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line\textbf{, notable for being detected at nearly all subset grid points}.
    Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.
    This and all spectral lines shown in this paper were identified using the NIST spectral line database \citep{NIST}.}
    \label{fig:figure3}
\end{figure*}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure4}
    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
    \textbf{This plot is much smaller than \autoref{fig:figure3} because} \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
    \label{fig:figure4}
\end{figure}

@@

Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator}, which reconstructs the spectrum by constructing a forward model based on the input state dictionary.
Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra.}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure5}
    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.
    We can see the spectral line shape be reconstructed mostly between the two native PHOENIX spectra.}
    \label{fig:figure5}
\end{figure}

We can observe in \autoref{fig:figure6} that \textbf{a Fe I line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure6}
    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel intrpolation of the 10123.66 \AA \ He II spectral line.}}
    \label{fig:figure6}
\end{figure}
It is interpolating the spectral line properties using hundreds of thousands of manifolds, each representing a nonlinear parameter in the shape of the spectral line.

@@

Performance results are shown in \autoref{fig:figure7}, and we can see that \textbf{the PHOENIX generator's parallel mode} is much faster than \textbf{serial mode}.
\textbf{While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
The speedup factor in parallel mode compared to KORG is then the same speedup factor from the serial mode.}

\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513 (32c/64t, 3.65 GHz boost)\\
        RAM & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{\textbf{Specifications of the} machine used for computation.
    \textbf{This machine is owned by Caroline Morley.}}
    \label{tab:table2}
\end{table}

@@

\subsection{Inference Algorithm}
\textbf{The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.}
We elected to use Bayesian optimization \textbf{for this component}, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{linenomath*}
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^2}^{1/2}
\end{gather}
\end{linenomath*}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
\textbf{A random evaluation is simply a randomly generated tuple of stellar parameters within the parameter space of the PHOENIX generator, which is then evaluated with the generator to get a reconstruction and then compared with the inference target to retrieve its RMS error.
The surrogate model is the approximation to the true objective function that the optimizer constructs based on the set of random evaluations (in this case defined by a set of 100 points rather than a fully continuous, infinitely resolved function)
Guided evaluations are different, they are not randomly generated tuples of stellar parameters.
They use the surrogate model in order to select new test points for the minimum of the objective function.}
This totals to 120 evaluations, \textbf{100} to create \textbf{the} surrogate model, then \textbf{20} guided evaluation\textbf{s}, which was deemed sufficient for this study.
Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
\textbf{Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.}
One inference run takes on average just under 7.5 minutes to complete.

@@

We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data \textbf{such as photometric color index}.
So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
$\log(g)$ and [Fe/H] were allowed to vary freely.
The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ values, and 2 unique [Fe/H] values, totaling 54 unique spectra in the test set.
$T_{\mathrm{eff}}$ ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6 in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.

The results of the inference algorithm are as follows: $T_{\mathrm{eff}}$ differed from the true result by an average of 185 K or 2.6\%.
$\log(g)$ differed by an average of 0.19 or 6.8\%. [Fe/H] differed by an average of 0.12 dex, which is 24\% of our search range.
From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].

@@

\section{Conclusion}
In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{([Fe/H]: [-0.5, 0] dex, $T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], $\lambda$: [8038, 12849] \AA)}.
We create state dictionaries for all spectra in the PHOENIX subset \textbf{using \texttt{blase}, recasting spectra into lists of Voigt profiles, and} lossily compress \textbf{the spectral data by a} factor of around 20. 
The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.
We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.
We introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
\textbf{Testing on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0] dex) yields} average absolute deviations from true values of 185 K in $T_\mathrm{eff}$, 0.19 in $\log(g)$, and 0.12 dex in [Fe/H].
In its current state, our algorithm operates on spectra within the PHOENIX subset parameter ranges in \autoref{tab:table1}, requiring that the spectra not contain noticeable Doppler shifting, rotational broadening, or other confounding factors.
The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.

>>>>

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
    We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
    \textbf{This} create\textbf{s} the PHOENIX generator, \textbf{enabling parallelized spectral synthesis}.
    Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24 dex, and [Fe/H]: 0.056 dex for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26 dex, and [Fe/H]: 0.16 dex for $T_\text{eff} \ge$ 7000 K}.
    \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
    This proof-of-concept \textbf{study} shows that semi-empirical HML is a viable alternative to traditional approaches \textbf{in spectroscopic inference}.
\end{abstract}

@@

\begin{table*}
    \hspace*{0.6cm}\begin{tabular}{llll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Interval} & \bf{Sampling}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex & \bf{N/A}\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex & \bf{0.5 dex}\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K & \bf{100 K until 7000 K, then 200 K}\\
        Surface Gravity & $\log(g)$ & [2, 6] dex & \bf{0.5 dex} \\
        Wavelength & $\lambda$ & [8038, 12849] \AA & $\mathbf{R = 500,000}$\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study.
    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}.
    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}.
    This subset is comprised of 1,314 individual spectra: 73 $T_{\mathrm{eff}}$ values, 9 $\log(g)$ values, and 2 [Fe/H] values.}
    \label{tab:table1}
\end{table*}

@@

Spectral lines were often only detected in some spectra from the PHOENIX subset.
In \autoref{fig:figure2}, we show that different grid points \textbf{show} differing counts of detected spectral lines.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of the PHOENIX subset.
    We can see that the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    \textbf{Remember} that from $T_{\mathrm{eff}} = 7000$ K onward, PHOENIX's sampling increment changes from 100 K to 200 K.
    \textbf{Due to some combination of computational artifacts from our line identification assumptions or innate behavior of PHOENIX grid spectra themselves, outliers such as the obvious one here at $T_{\mathrm{eff}} = 7000$ K and $\log(g) = 6$ dex appear.
    Further exploration into why this happens is not in the scope of this study; for inference we are concerned with the lines themselves, not overall grid point behavior.}}
    \label{fig:figure2}
\end{figure*}

@@

Examples of the appearance of missing sections in heatmaps where a line does not appear \textbf{can be seen} in \autoref{fig:figure3} and \autoref{fig:figure4}\textbf{, which show heatmaps of the 11617.66 \AA \ C I and the 10123.66 \AA \ He II lines (identified with NIST \citep{NIST}) varying over a solar metallicity slice of the PHOENIX subset stellar parameter space}.
Across the entire PHOENIX subset, \texttt{blase} detected 128,723 spectral lines \textbf{with unique $\mu$}.
\textbf{On average, a PHOENIX spectrum has 9,167 detected spectral lines, with the minimum being 252 ($T_\text{eff}$ = 12000 K, $\log(g)$ = 6 dex, [Fe/H] = -0.5 dex) and the maximum being 34,551 ($T_\text{eff}$ = 2300 K, $\log(g)$ = 3 dex, [Fe/H] = 0 dex).}
Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one \textbf{was} interpolated \textbf{to map stellar parameters to spectral line properties}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a \textbf{C I} spectral line\textbf{, notable for being detected at nearly all subset grid points}.
    Notice the missing chunk in the top left; \textbf{this could perhaps be due to the line center shifting to the point that \texttt{blase} identified the chunk as a different line entirely (notice the line center offset increasing in magnitude approaching the chunk border).}
    \textbf{Since} \texttt{blase} did not detect a spectral line here, we artificially populate those points with lines that have $\ln(a) = -1000$ \textbf{during interpolation}.}
    \label{fig:figure3}
\end{figure*}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure4}
    \caption{Heatmap\textbf{s} showing how $\ln(a)$, $\sigma$\textbf{, $\gamma$, and $\mu' - \mu$} vary over the PHOENIX subset slice at solar metallicity of a He II spectral line.
    \texttt{blase} detects this line at a select chunk of grid points, leading to \textbf{data only existing between $T_\text{eff}$ 3600 K and 6600 K}.}
    \label{fig:figure4}
\end{figure}

@@

Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator}, which reconstructs the spectrum by constructing a forward model based on the input state dictionary.
Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.
\textbf{In \autoref{fig:figure5}, we see a comparison between a PHOENIX generator solar spectrum and two nearby native PHOENIX grid point spectra at the 10123.66 \AA \ He II spectral line.}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure5}
    \caption{Plot of \textbf{the 10123.66 \AA \ He II} spectral line shown at the closest PHOENIX grid points to the Sun's stellar parameters \textbf{(solid lines)}, as well as a solar spectrum reconstructed with the PHOENIX generator \textbf{(dotted line)}.}
    \label{fig:figure5}
\end{figure}

We can observe in \autoref{fig:figure6} that \textbf{our He II line from a} reconstructed solar spectrum is not simply a pixel interpolation between the nearest \textbf{PHOENIX} grid point \textbf{spectra}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure6}
    \caption{\textbf{Comparison between the PHOENIX generator's line interpolation and a standard pixel interpolation of the 10123.66 \AA \ He II spectral line.
    Notice that line interpolation results in more symmetrical line wings due to the use of Voigt profiles.}}
    \label{fig:figure6}
\end{figure}
It is interpolating the spectral line properties using hundreds of thousands of manifolds, each representing a nonlinear parameter in the shape of the spectral line.

@@

Performance results are shown in \autoref{fig:figure7}, and we can see that \textbf{the PHOENIX generator's parallel mode} is much faster than \textbf{serial mode}.
\textbf{While the generator takes around 15 seconds per spectrum in serial mode, this drops to 1 second per spectrum in parallel mode as soon as 30 reconstructions.
The speedup factor between the two methods increases as more spectra are generated, reaching a factor of 20 at 60 reconstructions.
Assuming linear time scaling with wavelength range, according to KORG benchmarks \citep{korg}, which generate spectra over a 500 \AA \ wavelength range, the PHOENIX generator in serial mode is approximately as fast as KORG and $\sim$30 times faster than Turbospectrum \citep{turbospectrum}.
The speedup factor in parallel mode compared to KORG is then the same variable speedup factor from serial mode (20 times faster at 60 reconstructions).}

\textbf{The computer used for this test has specifications shown in \autoref{tab:table2} (note that spectral reconstruction does not currently leverage the GPU):}
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513 (32c/64t, 3.65 GHz boost)\\
        RAM & 256 GB\\
        GPU & Nvidia A100 [40GB] ($\times$2)\\
        \hline
    \end{tabular}
    \caption{\textbf{Specifications of the} machine used for \textbf{cloning and reconstruction testing}.
    \textbf{This machine is owned by Caroline Morley.}}
    \label{tab:table2}
\end{table}

@@

\subsection{Inference Algorithm}
\textbf{The goals of this study's inference algorithm are to infer the stellar parameters $T_\text{eff}$, $\log(g)$, and [Fe/H] by comparing a spectrum with our reconstructions and solving the optimization problem.}
We elected to use Bayesian optimization \textbf{for this component}, specifically the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}.
This algorithm uses a Gaussian Process to model the objective function, which in this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum $\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{linenomath*}
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^2}^{1/2}
\end{gather}
\end{linenomath*}
The optimizer was configured to first run 100 random evaluations to seed the surrogate model, then run 20 more evaluations now guided by the surrogate model.
\textbf{A random evaluation is simply a randomly generated tuple of stellar parameters within the parameter space of the PHOENIX generator, which is then evaluated with the generator to get a reconstruction and then compared with the inference target to retrieve its RMS error.
The surrogate model is the approximation to the true objective function that the optimizer constructs based on the set of random evaluations (in this case defined by a set of 100 points rather than a fully continuous, infinitely resolved function)
Guided evaluations are different, they are not randomly generated tuples of stellar parameters.
They use the surrogate model in order to select new test points for the minimum of the objective function.}
This totals to 120 evaluations, \textbf{100} to create \textbf{the} surrogate model, then \textbf{20} guided evaluation\textbf{s}, which was deemed sufficient for this study.
Fine-tuning these numbers is possible, but simply not warranted for a proof-of concept method.
\textbf{Tweaking these numbers could lead to a more accurate inference, however it would incur additional computational cost as well.}
One inference takes on average \textbf{200 seconds, using the computer with specifications in \autoref{tab:table3}}.

\textbf{Inference algorithms that output more information and produce posterior distributions such as \texttt{starfish} \citep{starfish} or MINESweeper \citep{minesweeper} take on the order of an hour to complete an inference on supercomputers.
\texttt{astroNN} \citep{leung2019} is a more direct comparison, and represents an extremely powerful inference tool that performs inference thousands of times faster thanks to its neural network architecture.}
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD Ryzen 9 7950X3D (16c/32t, 5.7 GHz boost)\\
        RAM & 32 GB\\
        GPU & Nvidia GeForce RTX 4080 SUPER [16GB]\\
        \hline
    \end{tabular}
    \caption{\textbf{Specifications of the machine used for general testing and for generating visualizations.}
    \textbf{This machine is owned by Sujay Shankar.}}
    \label{tab:table3}
\end{table}

@@

We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data \textbf{such as photometric color index}.
So when testing the inference algorithm, we limited its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the true value on either side.
$\log(g)$ and [Fe/H] were allowed to vary freely.
The test sample consisted of 21 $T_{\mathrm{eff}}$ values\textbf{: [3000, 11000] K in steps of 400 K}, 5 $\log(g)$ values\textbf{: [2, 6] dex in steps of 0.5 dex} , and 2 [Fe/H] values\textbf{: [-0.5, 0] dex in steps of 0.5 dex}, totaling \textbf{210} unique spectra in the test set.
\textbf{Two test passes were run, meaning that 420 inference tests were run in total.}

In \autoref{fig:figure8}, we show the residuals of the inferred stellar parameters with respect to the true values.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure8}
    \caption{\textbf{Residual plots for the GP minimization inference algorithm, all plotted with respect to $T_\text{eff}$ (values aggregated across $\log(g)$ and [Fe/H]).
    Notice the interesting change in behavior of all residuals at around $T_\text{eff} = 7000$ K.}}
    \label{fig:figure8}
\end{figure*}
\textbf{We see that there is a distinction between the algorithm's behavior above and below $T_{\mathrm{eff}} = 7000$ K, so we define low and high temperature regimes accordingly.
For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24 dex}, and [Fe/H] \textbf{0.056} dex \textbf{(standard deviations of 139 K, 0.40 dex, and 0.15 dex, respectively)}.
\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26 dex, and [Fe/H] 0.16 dex (standard deviations of 293 K, 0.32 dex, and 0.23 dex, respectively).}
From this, we can see that $T_{\mathrm{eff}}$ was the most accurately inferred parameter, followed by $\log(g)$, and then [Fe/H].

@@

\section{Conclusion}
In this study, we have presented a proof-of-concept algorithm that \textbf{deals with} a subset of the PHOENIX spectral model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex, $\lambda$: [8038, 12849] \AA)}.
We create state dictionaries for all spectra in the PHOENIX subset \textbf{using \texttt{blase}, recasting spectra into lists of Voigt profiles, and} lossily compress \textbf{the spectral data by a} factor of around 20. 
The\textbf{y are available on} Zenodo at \url{https://zenodo.org/records/11246174} \citep{zenodo}.

We create the PHOENIX generator and implement a performant spectral reconstruction algorithm, enabling anyone to create reconstructions of PHOENIX spectra with continuously valued stellar parameters.

We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff}$ < 7000) regimes.
In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.

In its current state, our algorithm requir\textbf{es} that the spectra not contain noticeable Doppler shifting, rotational broadening, \textbf{noise}, or other confounding factors.
The methods discussed here represent a step down a road not traveled in spectral inference, and have the potential to become more advanced in the future by fully utilizing the strengths of physics-informed machine learning.
